humanCdGF_predict_cfu

### 2016-10-24 code/HumanCdGF_predict_cfu.Rmd
include down to 1 otu

test <- data.frame(model = c('boruta',rep('reiterative_rf', 6),rep('one_mouse_per_cg',4)),
                   numOtus = c(length(boruta_otus), length(otus_used[[1]]),length(otus_used[[2]]),length(otus_used[[3]]),
                               length(otus_used[[4]]),length(otus_used[[5]]),length(otus_used[[6]]),30,27,6,2),
                   Rsq = c(tail(rf_boruta$rsq, 1), rf_rsq, 0.72, 0.71, 0.75, 0.71)
)
##### feature reduction #####
### boruta
set.seed(seed)
rf_features <- Boruta(log_cfu_med ~ ., data = cfu_shared_colonized.df, maxRuns = 2000)
boruta_otus <- names(rf_features$finalDecision)[rf_features$finalDecision == 'Confirmed']
set.seed(seed)
rf_boruta <- randomForest(log_cfu_med ~ ., data = cfu_shared_colonized.df[, c('log_cfu_med', boruta_otus)],
                           importance = TRUE, ntree = n_trees)
### above mean features, reiteratively reduce to 2 otus 
###### used mean because median results in including many low values in early OTU selection steps
rf_rsq <- c()
otus_used <- c()
rf_otus <- colnames(cfu_shared_colonized.df)[-1]
for(i in 1:100){
  if(length(rf_otus) > 2){
  set.seed(seed)
  rf <- randomForest(log_cfu_med ~ ., data = cfu_shared_colonized.df[, c('log_cfu_med', rf_otus)],
                     importance = TRUE, ntree = n_trees)
  rf_import_otus <- importance(rf)[,1]
  otus_used[[i]] <- rf_otus
  rf_otus <- rownames(data.frame(rf_import_otus[rf_import_otus >= mean(rf_import_otus)]))
  rf_rsq[i] <- tail(rf$rsq, 1)
  } else {
    set.seed(seed)
    rf <- randomForest(log_cfu_med ~ ., data = cfu_shared_colonized.df[, c('log_cfu_med', rf_otus)],
                       importance = TRUE, ntree = n_trees)
    otus_used[[i]] <- rf_otus
    rf_rsq[i] <- tail(rf$rsq, 1)
    stop('Only one OTU remains')
  }
}
#### select top otus from reiteration through RF
optimum_otus <- otus_used[[which(rf_rsq == max(rf_rsq))]]
set.seed(seed)
rf_all <- randomForest(log_cfu_med ~ ., data = cfu_shared_colonized.df[, c('log_cfu_med', optimum_otus)],
                           importance = TRUE, ntree = n_trees)

@@

om_rsq_rf <- c() # rsq of each rf model
om_rsq <- c() # cumulative rsq of all models
om_import_otus <- data.frame(otu = colnames(select(one_mouse_df, contains('Otu'))))
one_df <- c()
om_pred <- c()
for(i in 1:20){
    set.seed(i)
    train <- one_mouse_df %>% 
      group_by(cage_id) %>% 
      sample_n(1) %>% ungroup() 
    test <- one_mouse_df %>% 
      filter(!mouse_id %in% train$mouse_id)
    set.seed(seed)
    rf_om <- randomForest(log_cfu_med ~ ., data = select(train, -mouse_id, -cage_id),
                             importance = TRUE, ntree = n_trees)
    om_rsq_rf[i] <- tail(rf_om$rsq, 1)
    om_pred <- rbind(om_pred,
                     data.frame(select(test, mouse_id, cage_id, log_cfu_med),
                          predicted=predict(rf_om, test)))
    om_rsq[i] <- 1 - (sum((om_pred$log_cfu_med - om_pred$predicted)^2)/
               sum((om_pred$log_cfu_med - mean(om_pred$log_cfu_med))^2))
    import_df <- importance(rf_om)
    colnames(import_df)[1] <- paste(train$mouse_id, collapse = ' ')
    om_import_otus <- merge(om_import_otus, import_df, by.x = 'otu', by.y = 'row.names')
    om_import_otus <- select(om_import_otus, -IncNodePurity)
}

#####
test <- c() # test if any model has same train set
  for(i in 2:21){
  test[i-1] <- colnames(om_import_otus)[i] %in% colnames(om_import_otus)[-c(1,i)]
}
any(test)
#####

>>>>

test <- data.frame(model = c('boruta',rep('reiterative_rf', 6),rep('one_mouse_per_cg',4)),
                   numOtus = c(length(boruta_otus), length(otus_used[[1]]),length(otus_used[[2]]),length(otus_used[[3]]),
                               length(otus_used[[4]]),length(otus_used[[5]]),length(otus_used[[6]]),30,27,6,2),
                   Rsq = c(tail(rf_boruta$rsq, 1), rf_rsq, 0.72, 0.71, 0.75, 0.71)
)
##### feature reduction #####
### boruta
set.seed(seed)
rf_features <- Boruta(log_cfu_med ~ ., data = cfu_shared_colonized.df, maxRuns = 2000)
boruta_otus <- names(rf_features$finalDecision)[rf_features$finalDecision == 'Confirmed']
set.seed(seed)
rf_boruta <- randomForest(log_cfu_med ~ ., data = cfu_shared_colonized.df[, c('log_cfu_med', boruta_otus)],
                           importance = TRUE, ntree = n_trees)
### above mean features, reiteratively reduce to 2 otus 
###### used mean because median results in including many low values in early OTU selection steps
rf_rsq <- c()
otus_used <- c()
rf_otus <- colnames(cfu_shared_colonized.df)[-1]
for(i in 1:100){
  if(length(rf_otus) > 1){
  set.seed(seed)
  rf <- randomForest(log_cfu_med ~ ., data = cfu_shared_colonized.df[, c('log_cfu_med', rf_otus)],
                     importance = TRUE, ntree = n_trees)
  rf_import_otus <- importance(rf, type = 1)
  otus_used[[i]] <- rf_otus
  rf_otus <- rownames(data.frame(rf_import_otus[rf_import_otus >= mean(rf_import_otus)]))
  rf_rsq[i] <- tail(rf$rsq, 1)
  } else {
    set.seed(seed)
    rf <- randomForest(log_cfu_med ~ ., data = cfu_shared_colonized.df[, c('log_cfu_med', rf_otus)],
                       importance = TRUE, ntree = n_trees)
    otus_used[[i]] <- rf_otus
    rf_rsq[i] <- tail(rf$rsq, 1)
    stop('Only one OTU remains')
  }
}
#### select top otus from reiteration through RF
optimum_otus <- otus_used[[which(rf_rsq == max(rf_rsq))]]
set.seed(seed)
rf_all <- randomForest(log_cfu_med ~ ., data = cfu_shared_colonized.df[, c('log_cfu_med', optimum_otus)],
                           importance = TRUE, ntree = n_trees)

@@

om_rsq_rf <- c() # rsq of each rf model
om_rsq <- c() # cumulative rsq of all models
om_import_otus <- data.frame(otu = colnames(select(one_mouse_df, contains('Otu'))))
one_df <- c()
om_pred <- c()
for(i in 1:20){
    set.seed(i)
    train <- one_mouse_df %>% 
      group_by(cage_id) %>% 
      sample_n(1) %>% ungroup() 
    test <- one_mouse_df %>% 
      filter(!mouse_id %in% train$mouse_id)
    set.seed(seed)
    rf_om <- randomForest(log_cfu_med ~ ., data = select(train, -mouse_id, -cage_id),
                             importance = TRUE, keep.forest = T, ntree = n_trees)
    om_rsq_rf[i] <- tail(rf_om$rsq, 1)
    om_pred <- rbind(om_pred,
                     data.frame(select(test, mouse_id, cage_id, log_cfu_med),
                          predicted=predict(rf_om, test)))
    om_rsq[i] <- 1 - (sum((om_pred$log_cfu_med - om_pred$predicted)^2)/
               sum((om_pred$log_cfu_med - mean(om_pred$log_cfu_med))^2))
    import_df <- importance(rf_om)
    colnames(import_df)[1] <- paste(train$mouse_id, collapse = ' ')
    om_import_otus <- merge(om_import_otus, import_df, by.x = 'otu', by.y = 'row.names')
    om_import_otus <- select(om_import_otus, -IncNodePurity)
}
plot(om_rsq_rf)
#####
test <- c() # test if any model has same train set
  for(i in 2:21){
  test[i-1] <- colnames(om_import_otus)[i] %in% colnames(om_import_otus)[-c(1,i)]
}
any(test)
#####
